---
title: "Mock Manuscript: Determination of pKa Using Spectrophotometric Speciation Analysis"
author: "Emiko Yesaki"
date: "2026-02-11"
output:
  pdf_document: default
  html_document: default
bibliography: references.bib
cls: peerj.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Mock Manuscript**

## **Title**

**Determination of pKa Using Spectrophotometric Speciation Analysis**

## **Author(s)**

E. Yesaki, University of British Columbia Okanagan

## **Abstract**

This study investigates the acid–base behavior of a pH indicator dye by measuring absorbance at two wavelengths (435 nm and 593 nm) across a range of buffer solutions. Using Beer–Lambert–derived concentrations of protonated and deprotonated species, we constructed speciation plots and applied the Henderson–Hasselbalch equation to estimate the indicator’s pKa. The two analytical approaches produced consistent pKa values, demonstrating the reliability of spectrophotometric methods for acid–base characterization.

## **Introduction**

Indicators undergo a color change as they transition between protonated and deprotonated forms. The pKa of an indicator defines the pH at which these species are present in equal concentrations. Spectrophotometry provides a convenient method for monitoring this transition because each species absorbs light at different wavelengths. In this experiment, absorbance measurements were used to quantify species distributions and estimate the pKa of the indicator dye.

## **Methods**

### **Sample Preparation**

Buffer solutions spanning a range of pH values were prepared according to the Lab T protocol. A fixed concentration of indicator dye was added to each solution to ensure consistent absorbance measurements.

### **Sample Running**

Absorbance at 435 nm and 593 nm was measured for each sample using a calibrated spectrophotometer. pH was recorded using a standard pH meter. All raw measurements were stored in data_raw_2025-01-26.csv.

### **Data Analysis**

Absorbance values were converted to concentrations using the Beer–Lambert law and known molar absorptivities. The ratio $[In^{2-}]/[HIn^{-}]$ and its logarithm were calculated. pKa was estimated using: 1. **Speciation midpoint** (ratio = 1)\
2. **Henderson–Hasselbalch regression** of pH vs. log ratio

Processed values were saved in lab-t_data_processed_2026-02-12.csv.

## **Results**

The indicator exhibited a clear transition between protonated and deprotonated forms across the measured pH range. Speciation analysis produced a pKa of 8.78, while the Henderson–Hasselbalch plot yielded a pKa of 8.12. The close agreement between methods supports the internal consistency of the dataset.

## **Discussion**

Both analytical approaches produced similar pKa values, indicating that spectrophotometric measurements reliably capture the acid–base behavior of the indicator. Minor differences between methods likely reflect experimental noise, pH meter precision, and wavelength‑specific sensitivity. These results align with expected indicator behavior and demonstrate the usefulness of spectrophotometric speciation analysis in undergraduate laboratory settings. This is a test citation @zhang2022. Another citation @brito2024.

## **Conclusion**

```{r echo=FALSE}
knitr::kable(
  data.frame(
    Method = c("Speciation midpoint", "Henderson–Hasselbalch regression"),
    Estimated_pKa = c("pKa_speciation", "pKa_HH"),
    Description = c(
      "pH at which [In2-]/[HIn-] = 1",
      "Intercept of linear regression of pH vs log ratio"
    )
  ),
  caption = "Comparison of pKa estimation methods used in Lab T."
)
```

## References

::: {#refs}
:::

